---
title: "Transcriptomics-Seurat Workflow-GSE111672"
output: html_notebook
---
Daniela Q. 
Transcriptomics Course

Dataset: GSE111672.
Pancreatic ductal adenocarcinoma. This analysis focuses on one of the two samples in this study 

Install Necessary Packages

```{r}
install.packages('patchwork')
```

```{r}
library(dplyr)
library(Seurat)
library(patchwork)
```
```{r}
library(ggplot2)
```



Load the TSV file

```{r}
data=read.table(file = '/Users/danielaquijano/Documents/GitHub/Transcriptomics-Course-Projects/source_files/GSM3036909.tsv', sep = '\t', header = TRUE)
```

```{r}
PDAC.seurat.obj <- CreateSeuratObject(counts = data, project = "PDAC", min.cells = 3, min.features = 200)
```
Label Mitochondrial genes
```{r}
PDAC.seurat.obj[["percent.mt"]] <- PercentageFeatureSet(PDAC.seurat.obj, pattern = "^MT-")
head(PDAC.seurat.obj@meta.data)
```
View the number of counts, number of features, and the percent mitochondrial genes.
```{r}
VlnPlot(PDAC.seurat.obj, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
```
Keep the cells that have greater than 200 and less than 2500 unique features and the percent mitochondrial genes is less than 5.

```{r}
PDAC.seurat.obj <- subset(PDAC.seurat.obj, subset = nFeature_RNA > 200 & nFeature_RNA < 2500 &  percent.mt < 5)
```

Normalize data using the LogNormaliza method with a scale factor of 10,000
```{r}
#Normalize Data
PDAC.seurat.obj  <- NormalizeData(
  PDAC.seurat.obj,
  normalization.method = "LogNormalize",
  scale.factor = 10000,
  verbose = TRUE, )
str(PDAC.seurat.obj)
```

Find the 2000 most variable genes using the FindVariableFeatures command using the vst method.
```{r}
PDAC.seurat.obj <- FindVariableFeatures(PDAC.seurat.obj, selection.method = "vst", nfeatures = 2000)
```

Scale data to normalize the standard deviation and center the data.
```{r}
all.genes <- rownames(PDAC.seurat.obj)
PDAC.seurat.obj <- ScaleData(PDAC.seurat.obj, features = all.genes)
```

Perform Principal Component Analysis

Based on the PCA, it seems that there are two main clusters of data (based on DimPlot)
```{r}
PDAC.seurat.obj <- RunPCA(PDAC.seurat.obj, features = VariableFeatures(object = PDAC.seurat.obj))
```
```{r}
# visualize PCA results
print(PDAC.seurat.obj[["pca"]], dims = 1:10)
```
```{r}
for (i in 1:10) {
  DimHeatmap(PDAC.seurat.obj, dims = i, cells = 500, balanced = TRUE)
}

```



Visualize data using VizDimLoadings and DimPlot functions. Can you tell from the PCA analyis, the number of cell types that are present?
```{r}
VizDimLoadings(PDAC.seurat.obj, dims = 1:2, reduction = "pca")
```

```{r}
VizDimLoadings(PDAC.seurat.obj, dims = 3:4, reduction = "pca")
```
```{r}
VizDimLoadings(PDAC.seurat.obj, dims = 5:6, reduction = "pca")
```
```{r}
VizDimLoadings(PDAC.seurat.obj, dims = 7:8, reduction = "pca")
```
```{r}
VizDimLoadings(PDAC.seurat.obj, dims = 9:10, reduction = "pca")
```
```{r}
DimPlot(PDAC.seurat.obj, reduction = "pca")
```

Find when does the variation reaches the lowest amount of variation explained. Using the JackStraw method with 100 replicates and score the first 20 dimensions.

```{r}
PDAC.seurat.obj <- JackStraw(PDAC.seurat.obj, num.replicate = 100)
PDAC.seurat.obj <- ScoreJackStraw(PDAC.seurat.obj, dims = 1:20)
```

Plot the results for the first 20 dimensions.

```{r}
JackStrawPlot(PDAC.seurat.obj, dims = 1:20)

```

Use the elbow plot
```{r}
ElbowPlot(PDAC.seurat.obj)
```
Group together the cells based on where they are located in the different dimensions. Use the FindNeighbors function using the first 9 dimensions.

```{r}
# 7. Clustering ------------
PDAC.seurat.obj<- FindNeighbors(PDAC.seurat.obj, dims = 1:9)
```

Identify the clusters usng the FindClusters function.

```{r}
PDAC.seurat.obj <- FindClusters(PDAC.seurat.obj)
head(PDAC.seurat.obj@meta.data)
```

Perform a UMAP analysis using the first 9 dimensions using RunUMAP and then visualize it using DimPlot.

Here UJMAP was done on the fist 9 dimensions (dims=1:9) and there are 9 clusters assigned in total. There are some samples in clsuter 6 that seem to not belong and some ambiguously assigned samples in between clusters 7 and 2 and in the top right 'cluster of clusters' there seems to be some ambiguous assignment as well. 

```{r}
PDAC.seurat.obj <- RunUMAP(PDAC.seurat.obj, dims = 1:9)

DimPlot(PDAC.seurat.obj, reduction = "umap")
```

Identify the markers that compare each cluster against all. Report only positively markers (only.pos=TRUE). Using FindAllMarkers.

```{r}
PDAC.markers <- FindAllMarkers(PDAC.seurat.obj, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
PDAC.markers %>% group_by(cluster) %>% top_n(n = 2, wt = avg_log2FC)
```



Create a violin plot using one feature from each cluster.

```{r}
genes_list = c("16289", "10632","4231","6649", "12925","3135","7771", "2136")

```

```{r}
VlnPlot(PDAC.seurat.obj, features=c("16289", "10632","4231"))
```

```{r}
VlnPlot(PDAC.seurat.obj, features=c("6649", "12925","3135"))
```
```{r}
VlnPlot(PDAC.seurat.obj, features=c("7771", "2136"))
```



Create a feature plot using the same features as before.

```{r}
FeaturePlot(PDAC.seurat.obj, features=c("16289", "10632","4231"))

```
```{r}
FeaturePlot(PDAC.seurat.obj, features=c("6649", "12925","3135"))
```
```{r}
FeaturePlot(PDAC.seurat.obj, features=c("7771", "2136"))
```


